---
title: "GSS challange Less Data"
author: "Cynthia Cornejo"
date: "July 3. 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```

The white race is more supportive of the death epnalty than the other race, which is second in support behind the black race. With the other race coming in second, there are more members of the black race who oppose the death penalty.

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The majority of white people (67.75%) still support the death penalty, with the other races having a ratio of 61.25 that is extremely close to that of the white people. Of all racial groups, 52.76 percent of black people oppose the death penalty. While still having an equal number of members in each race. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

With a p-value of .005, there is still a difference in the statistics. 

```{r}
fisherXY(mydata$race,mydata$cappun, printAll = TRUE)
```

Black and white people are different races. Both do not differ from one another. Even though there was less data, the white and black races have the only significant p-value. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

According to this data, the biggest percentage of supporters of the death penalty still belong to the white race, while the highest proportion of opponents belongs to the black race. Based on the data presented and their p-values, the only distinction between the black and white racial groups is made in this set of graphs and statistics. So yes, there is a difference between the number two branch and the master, and the explanation for this is given above. However, in terms of support for the death penalty throughout both branches, white people continue to hold the majority. 

